Object Capability Patterns:\\Policies and Specifications

### 2017-08-16 article.tex
update

\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{multirow}
\usepackage{multicol}
\usepackage[left=2.5cm,right=2cm,top=2cm,bottom=2cm, hmarginratio=1:1,headheight=14pt,columnsep=15pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{graphicx}
\usepackage{enumitem} % Customized lists

\setlist[itemize]{noitemsep} % Make itemize lists more compact

@@

%----------------------------------------------------------------------------------------
%	Custom Logic Environment and  Bindings
%----------------------------------------------------------------------------------------
%eq numbering
\makeatletter
\newcommand{\eqnum}[1][]{\leavevmode\hfill\refstepcounter{equation}\label{#1}\textup{\tagform@{\theequation}}}
\makeatother
%environment formating
\newenvironment{logic}[1][null]
{\begin{flushleft} \small \label{#1}}
{\end{flushleft}}
%logic shortcuts
\newcommand*\rot{\rotatebox{90}}
\newcommand{\loin}{$\in$}
\newcommand{\loforall}{$\forall$}
\newcommand{\loexists}{$\exists$}
\newcommand{\loand}{$\land$}
\newcommand{\loor} {$\lor$}
\newcommand{\loneq} {$\neq$}
\newcommand{\losimeq} {$\simeq$}
\newcommand{\losubseteq}{$\subseteq$}
\newcommand{\loimplies}{$\implies$}
\newcommand{\lonimplies}{$\centernot\implies$}
\newcommand{\loimpliedby}{$\Longleftarrow$}
\newcommand{\lonimpliedby}{$\centernot\Longleftarrow$}
\newcommand{\losigma}{$\upsigma$}
\newcommand{\loturns} {$\vDash$}
\newcommand{\lochi}{$\upchi$}
\newcommand{\lophi}{$\upvarphi$}
\newcommand{\lorarrow}{$\rightarrow$}
\newcommand{\lonturns} {$\nvDash$}
\newcommand{\loiff} {$\iff$}
\newcommand{\loleadsto} {$\leadsto$}
\newcommand{\lomapsto} {$\mapsto$}
\newcommand{\locup} {$\cup$}
\newcommand{\lobigcup} {$\bigcup$}
\newcommand{\lotimes} {$\times$}
\newcommand{\loneg}{$\boldsymbol \neg$}
\newcommand{\loexec}[2] {$\lfloor$#1$\rfloor _{\text{#2}}$}
\newcommand{\loconj}[1] {$\bar{\text{#1}}$}
\newcommand{\ablock} {\null\qquad}

@@

With these simplifications, protection for us then becomes solely a matter of whether we can allow or deny an object to modify the state of another. In the following \ref{protection} below we build the neccessary conditions for MayAffect$^{Eve}$, where MayAffect$^{Eve}$ is placed in the antecedent, and we examine which of the family of MayAccess predicates is placed in the consequent. To help us achieve this, we also formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions in \ref{ocaprules}.

@@

\textbf{Rule 1: Objects can only interact with each other through sending messages on capabilities.}
We begin our reasoning of protection in our OCap model, by first re-asserting the necessary but not sufficient path condition of capability, and calling it \textsc{pec}. By definition of our predicates, an object having a capability of o' implies object o having a path to o', but not vice-versa:
\begin{logic}
\hrule\null
\textsc{\normalsize *Path-Execution Connectivity (PEC)}\\
M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(o,o') \loimplies \linebreak
\ablock \ablock  MayAccess$^{Ind,Now}$(o,o')
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ \loneg MayAccess$^{Indir Now}$(o,o')\loimplies \linebreak
\ablock \ablock \loneg MayAccess$^{Dir,Now}$(o,o')
\linebreak
\hrule
\end{logic}
Note also that \textsc{pec} is a direct intepretation from Rule 1, since rule 1 says owning a capability is the only way of sending messages to another object, and therefore implies a path between the objects.\\

\textbf{Rule 2: Objects cannot forge capabilities, and only connectivity begets connectivity.}
\begin{logic}
\hrule\null
\textsc{\normalsize *Global Path Existence (GPE)}\\
M,\losigma\ \loturns\ $[$\loexists Y*\loin{Obj}. MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
\ablock \loiff\linebreak
\ablock $[$\loexists X*\loin{Obj}. MayAccess$^{Dir,Now}$(X*,o')$]$
\linebreak\\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
\ablock \loiff\linebreak
\ablock $[$\loforall Y*\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
\hrule
\end{logic}
In addition to path being a necessary condition for capability, \textsc{gpe} gives us an additional new relation between the two concepts over the entire system. \textsc{gpe} says that in a given system of objects, $iff$ there exists an object Y* which has a path to o', then there exists an object X* that has the capability of o'. Looking from the left to right direction, since the path from Y* to o' exists, then there must exist an object X* with a capability of o' so that the path from Y* to o' is well-established. Moving from the right to left direction, \textsc{gpe} says that if an object X* has a capability then there exist an object Y* with a path to o', and this can be proven from \textsc{pec} or by definition when (X*==Y*).\\

However, \textsc{gpe} only postulates the existence of some object with some capability iff there exists some object with some path, and says nothing about how the objects are connected. From the connectivity begets connectivity rule, \textsc{pec}, and \textsc{gpe}, we construct global path-execution and execution-execution relationships over the entire system:
\begin{logic}
\hrule\null
\textsc{\normalsize *Global Path-Execution Connect. (GPEC)}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(o,o')\linebreak
\ablock \loiff\linebreak
\ablock \loexists X*\loin{Obj}. $[$MayAccess$^{Ind,Now}$(o, X*) \loand\linebreak
\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o')$]$
\linebreak\\
\textit{By contraposition,} \\
M,\losigma\ \loforall X*\loin{Obj}. \loneg $[$MayAccess$^{Ind,Now}$(o, X*) \loand\linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
\ablock \loiff\linebreak
\ablock \loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
\hrule
\end{logic}
\textsc{gpec} introduces an intermediate object X* between o and o' that makes explicit the missing connection in \textsc{gpe}.\\
The power in \textsc{gpec} lies in the recursive predicate MayAccess$^{Ind,Now}$(o, X*) that can eventually be expanded into a set of MayAccess$^{Dir,Now}$ predicates that finishes with terminating predicate MayAccess$^{Ind,Now}$(o, o), where the terminating predicate will always be true.
We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations arising from \losigma\ ($Eve$).
\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Connectivity 1 (EPC1)}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Ind,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
\ablock \loimplies\linebreak
	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
\hrule
\end{logic}
\textsc{epc1} has the meaning that if an object o has an eventual path to o' in some arising state \losigma', then there must exist an object X* that has an eventual path to o', and there must exist a way for o to have a path to X* in \losigma'. o either must already have a path to X* in \losigma, or that it will eventually receive the capability of X* in \textsc{epc1} is a representation of connectivity begets connectivity across time:

\begin{itemize}
\item \textbf{Initial Conditions or Endowment:} o has an existing path to o' in \losigma, therefore X* refers to o'
\item \textbf{Parenthood:} o will create o' in some arising \losigma', therefore X* refers to o
\item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma', therefore X* refers to an object that is not the same object as o (X*\loneq o). 
\end{itemize}
Note that for o to have an eventual path to o', we require only o to have an eventual path to X* since we have stated that X* will have an eventual path to o'. If o already has a path to X* in \losigma\ then we know o can reach X* in \losigma'. If not, the capability of X* must be introduced to o. For X* to introduce itself, X* must have the capability of o in \losigma'. Well, what if the capability of X* is introduced by some \textit{other} object \~X*? Note that in such a case, \~X* will have the capability of X*, and will therefore also be able to eventually have a path to o'. Also \~X* must also be able to introduce itself to o. There is hence no logical difference between \~X* and X* in our formal description and \~X* might simply be referred to as X*.\\

The contraposition result of \textsc{epc1} says there can be no eventual paths from o to o'. This can be achieved from:
\begin{itemize}
\item Global two-way path isolation of o in present state \losigma, where there are no paths leading to o \textit{and} no paths originating from o in \losigma
\item Global one-way path isolation to o' in all possible states \losigma'\end{itemize}
There is a final critical result from \textsc{epc1}.
Notice how, there is a 'recursive' MayAccess$^{Ind,Eve}$(X*,o') in our condition for MayAccess$^{Ind,Eve}$(o,o').
This allows us to recursively 'expand' the condition to incorporate \textit{all} X* intermediate objects in the path leading to o' that gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true. This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration completely based in \losigma.\\
What about capabilities? Using \textsc{gpec}, we can always expand MayAccess$^{Ind, Now}$(o,X*) into a set of MayAccess$^{Dir, Now}$ predicates that terminate with MayAccess$^{Ind, Now}$(o,o) which is always true. The final result we will get is a configuration based purely on MayAccess$^{Dir, Now}$ predicates!!!

\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Connectivity 2 (EPC2)}\\
M,\losigma\ \loturns\ 
MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies \linebreak
\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock  \ablock MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \quad \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
\ablock \ablock \ablock \ablock  $\}$\linebreak
\linebreak \\
\textit{By contraposition and De Morgan rules}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \loneg \loexists Y*\loin Obj. $[$MayAccess$^{Ind,Eve}$(X*,Y*) \loand\linebreak
\ablock \ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(Y*,o')$]$\linebreak
\ablock \ablock \ablock \ablock  $\}$\linebreak
\ablock \loimplies\linebreak
\ablock \loneg MayAccess$^{Ind,Eve}$(o,o')\
\linebreak
\hrule
\end{logic}
The results of \textsc{epc2} are identical to \textsc{epc1}, with the second condition of \textsc{epc1} expanded in \textsc{epc2} to say global one-way path isolation to o in all possible states \losigma' to mean there does not exist any eventual path from any object to an object Y* that has the eventual capability of o'.\\

So far we have developed results for paths, but what about execution? We construct \textsc{eec} from \textsc{epc1}, the only difference being we now require all those paths in \textsc{epc1} to be traversable.

\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Execution Connectivity (EEC)}\\
M,\losigma\ \loturns\ MayAccess$^{Dir,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Dir,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \ablock MayAccess$^{Dir,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Dir,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg $[$MayAccess$^{Dir,Eve}$(X*,o')$]\}$\linebreak
\ablock \loimplies\linebreak
	\ablock\loneg MayAccess$^{Dir,Eve}(o,o')$\linebreak
\hrule
\end{logic}
The power of \textsc{eec} is similar to \textsc{epc}\ in that it allows an arbitrary recursive expansion of the MayAccess$^{Dir,Eve}$(X*,o') till the terminating MayAccess$^{Dir,Eve}$(o',o') that is always true.

We now turn our attention formalzing Ocap state changes rules in the system.\\

\textbf{Rule 3: Objects may have private encapsulation of state and behaviour}
\begin{logic}
\hrule\null
\textsc{\normalsize *Global State Change Existence (GSCE)}\\
M,\losigma\ \loturns\ $[$\loexists X*\loin{Obj}. MayAffect$^{Now}$(X*,o')$]$\linebreak
\ablock \loimplies \linebreak
\ablock $[$\loexists Y*\loin{Obj}. MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
\ablock \loimplies \linebreak
\ablock $[$\loexists Z*\loin{Obj}. MayAccess$^{Ind,Now}$(Z*,o')$]$
\linebreak\\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ $[$\loforall Z*\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(Z*,o')$]$\linebreak
\ablock \loimplies\linebreak
\ablock $[$\loforall Y*\loin{Obj}. \loneg MayAccess$^{Dir,Now}$(Y*,o')$]$\linebreak
\ablock \loimplies\linebreak
\ablock $[$\loforall X*\loin{Obj}. \loneg MayAffect$^{Dir,Now}$(X*,o')$]$\linebreak
\hrule
\end{logic}
\textsc{gsce} can be derived from our predicate definitions. The first implication states that if an object can change the state of o', then there must exist an object in the system that holds the capability of o'. The second implication is that there must exist an object in the system that holds a path to o'.
We view reasoning about state change as reasoning about protection, which we elaborate in the next subsection.
\subsection{Protection}\label{protection}
To reason about enforcing protection, we focus on finding the necessary conditions for the weaker predicate MayAffect$^{Eve}$ rather than MayAffect$^{Now}$ in the antecedent. This is because negation on both sides of the implication, would yield a stronger \loneg MayAffect$^{Eve}$ in the consequent, which is our area of interest (since it represents state protection). In practical terms, if we are concerned with the protection of o' from o, it is also often not very useful to have a policy where \loneg MayAffect$^{Now}$(o,o') holds but \loneg MayAffect$^{Eve}$(o,o') does not. Furthermore, by ensuring \loneg MayAffect$^{Eve}$(o,o') holds, we can also ensure \loneg MayAffect$^{Now}$(o,o') holds since by contraposition:

>>>>

\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage{multirow}
\usepackage{multicol}
\usepackage[left=2.5cm,right=2cm,top=2cm,bottom=2cm, hmarginratio=1:1,headheight=14pt,columnsep=15pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{graphicx}
\usepackage{enumitem} % Customized lists
\usepackage{color}
\setlist[itemize]{noitemsep} % Make itemize lists more compact

@@

%----------------------------------------------------------------------------------------
%	Custom Logic Environment and  Bindings
%----------------------------------------------------------------------------------------
%eq numbering
\makeatletter
\newcommand{\eqnum}[1][]{\leavevmode\hfill\refstepcounter{equation}\label{#1}\textup{\tagform@{\theequation}}}
\makeatother
%environment formating
\newenvironment{logic}[1][]
{\begin{flushleft} \small }
{\end{flushleft}}
%logic shortcuts
\newcommand*\rot{\rotatebox{90}}
\newcommand{\loin}{$\in$}
\newcommand{\loforall}{$\forall$}
\newcommand{\loexists}{$\exists$}
\newcommand{\loand}{$\land$}
\newcommand{\loor} {$\lor$}
\newcommand{\loneq} {$\neq$}
\newcommand{\losimeq} {$\simeq$}
\newcommand{\losubseteq}{$\subseteq$}
\newcommand{\loimplies}{$\implies$}
\newcommand{\lonimplies}{$\centernot\implies$}
\newcommand{\loimpliedby}{$\Longleftarrow$}
\newcommand{\lonimpliedby}{$\centernot\Longleftarrow$}
\newcommand{\losigma}{$\upsigma$}
\newcommand{\loturns} {$\vDash$}
\newcommand{\lochi}{$\upchi$}
\newcommand{\lophi}{$\upvarphi$}
\newcommand{\lorarrow}{$\rightarrow$}
\newcommand{\lonturns} {$\nvDash$}
\newcommand{\loiff} {$\iff$}
\newcommand{\loleadsto} {$\leadsto$}
\newcommand{\lomapsto} {$\mapsto$}
\newcommand{\locup} {$\cup$}
\newcommand{\lobigcup} {$\bigcup$}
\newcommand{\lotimes} {$\times$}
\newcommand{\loneg}{$\boldsymbol \neg$}
\newcommand{\loexec}[2] {$\lfloor$#1$\rfloor _{\text{#2}}$}
\newcommand{\loconj}[1] {$\bar{\text{#1}}$}
\newcommand{\ablock} {\null\qquad}

@@

With these simplifications, protection for us then becomes solely a matter of whether we can allow or deny an object to modify the state of another. To help us reason about protection, we first formalize our assumptions and the rules of OCap, using our predicates, to help guide us in constructing our necessary conditions in \ref{ocaprules}. In the following \ref{protection} below we build the neccessary conditions for MayAffect$^{Eve}$, where MayAffect$^{Eve}$ is placed in the antecedent, and we examine which of the family of MayAccess predicates is placed in the consequent.

@@

\textbf{Rule 1: Objects can only interact with each other through sending messages on capabilities.}
We begin our reasoning of protection in our OCap model, by first re-asserting the necessary but not sufficient path condition of capability, and calling it \textsc{pec}. By definition of our predicates, an object having a capability of o' implies object o having a path to o', but not vice-versa:
\begin{logic}
\hrule\null
\textsc{\normalsize *Path-Execution Connectivity (PEC)}\\
M,\losigma\ \loturns\ MayAccess$^{Dir,Now}$(o,o') \loimplies \linebreak
\ablock \ablock  MayAccess$^{Ind,Now}$(o,o')
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ \loneg MayAccess$^{Ind,Now}$(o,o')\loimplies \linebreak
\ablock \ablock \loneg MayAccess$^{Dir,Now}$(o,o')
\linebreak
\hrule
\end{logic}
Note also that \textsc{pec} is a direct intepretation from Rule 1, since rule 1 says owning a capability is the only way of sending messages to another object, and therefore implies a path between the objects.\\

\textbf{Rule 2: Objects cannot forge capabilities, and only connectivity begets connectivity.}
\begin{logic}
\hrule\null
\textsc{\normalsize *Global Path Existence (GPE)}\\
M,\losigma\ \loturns\ $[$\loexists Y*\loin{Obj}. MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
\ablock \loiff\linebreak
\ablock $[$\loexists X*\loin{Obj}. MayAccess$^{Dir,Now}$(X*,o')$]$
\linebreak\\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ $[$\loforall X*\loin{Obj}. \loneg MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
\ablock \loiff\linebreak
\ablock $[$\loforall Y*\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(Y*,o')$]$\linebreak
\hrule
\end{logic}
In addition to path being a necessary condition for capability, \textsc{gpe} gives us an additional new relation between the two concepts over the entire system. \textsc{gpe} says that in a given system of objects, $iff$ there exists an object Y* which has a path to o', then there exists an object X* that has the capability of o'. Looking from the left to right direction, since the path from Y* to o' exists, then there must exist an object X* with a capability of o' so that the path from Y* to o' is well-established. Moving from the right to left direction, \textsc{gpe} says that if an object X* has a capability then there exist an object Y* with a path to o', and this can be proven from \textsc{pec} or by definition when (X*==Y*).\\
The contraposition result of \textsc{gpe} says that we can prevent all paths to o' through one of these conditions:
\begin{itemize}
\item Deny all direct paths leading into o' 
\end{itemize}

However, \textsc{gpe} only postulates the existence of some object with some capability iff there exists some object with some path, and says nothing about how the objects are connected. From the connectivity begets connectivity rule, \textsc{pec}, and \textsc{gpe}, we construct global path-execution and execution-execution relationships over the entire system:
\begin{logic}
\hrule\null
\textsc{\normalsize *Global Path-Execution Connect. (GPEC)}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Now}$(o,o')\linebreak
\ablock \loiff\linebreak
\ablock \loexists X*\loin{Obj}. $[$MayAccess$^{Ind,Now}$(o, X*) \loand\linebreak
\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o')$]$
\linebreak\\
\textit{By contraposition,} \\
M,\losigma\ \loforall X*\loin{Obj}. $[$\loneg MayAccess$^{Ind,Now}$(o, X*) \loor\linebreak
\ablock \ablock \ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o')$]$\linebreak
\ablock \loiff\linebreak
\ablock \loneg MayAccess$^{Ind,Now}$(o,o')\linebreak
\hrule
\end{logic}
\textsc{gpec} introduces an intermediate object X* between o and o' that makes explicit the missing connection in \textsc{gpe}.\\
The contraposition result of \textsc{gpec} says that we can prevent all paths from o to o' through one of these conditions:

The contraposition result of \textsc{gpec} says that for there to be no paths from o to o' these configurations must hold:
\begin{itemize}\setlength\itemsep{0.5em}
\item if there is some object x that o has a path to in \losigma, then x cannot have a path to o' in \losigma.
\item if there is some object x that has a path to o' in \losigma, then o cannot have a path to x in \losigma.
\end{itemize}

The power in \textsc{gpec} lies in the recursive predicate MayAccess$^{Ind,Now}$(o, X*) that can eventually be expanded into a set of MayAccess$^{Dir,Now}$ predicates that finishes with terminating predicate MayAccess$^{Ind,Now}$(o, o), where the terminating predicate will always be true, unless the object o does not exist in \losigma.
We now look for a relation between a path configuration at a state \losigma\ ($Now$) and eventual path configurations in states \losigma' arising from \losigma\ ($Eve$).\vfill
\begin{minipage}[b]{\linewidth}
\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Connectivity 1 (EPC1)}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj. $\{[$MayAccess$^{Ind,Now}$(o,X*)\linebreak
\ablock \ablock \ablock \ablock\loor \linebreak
\ablock \ablock \ablock \quad\: MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \quad \loand \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$\}$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ \linebreak
\ablock \loforall X*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(o,X*)\linebreak
\ablock \ablock \ablock \ablock \quad \loand \linebreak
\ablock \ablock \ablock \quad\: \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \quad \loor \linebreak
\ablock \ablock \ablock \quad \loneg MayAccess$^{Ind,Eve}$(X*,o')$\}$\linebreak
\ablock \loimplies\linebreak
	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
\hrule
\end{logic}
\end{minipage}
\textsc{epc1} has the meaning that if an object o has an eventual path to o' in some arising state \losigma', then there must exist an object X* that has an eventual path to o', and there must exist a way for o to have a path to X* in \losigma'. o either must already have a path to X* in \losigma, or that it will eventually receive the capability of X* in \losigma'. \textsc{epc1} is actually a formal representation of connectivity begets connectivity across time:

\begin{itemize}
\item \textbf{Initial Conditions or Endowment:} o has an existing path to o' in \losigma, therefore X* refers to o'
\item \textbf{Parenthood:} o will create o' in some arising \losigma', therefore X* refers to o
\item \textbf{Introduction:} o will only obtain the path to o' in some arising \losigma', therefore X* refers to an object that is not the same object as o (X*\loneq o). 
\end{itemize}

Note that for o to have an eventual path to o', we require only o to have an eventual path to X* since we have stated that X* will have an eventual path to o'. If o already has a path to X* in \losigma\ then we know o can reach X* in \losigma'. If not, the capability of X* must be introduced to o. For X* to introduce itself, X* must have the capability of o in \losigma', and the only way we can guarantee is to have X* possess the capability of o in \losigma. Well, what if the capability of X* is introduced by some \textit{other} object \~X*? Note that in such a case, \~X* will have the capability of X*, and will therefore also be able to eventually have a path to o'. Also \~X* must also be able to introduce itself to o. There is hence no logical difference between \~X* and X* in our formal description and \~X* might simply be referred to as X*.\\

The contraposition result of \textsc{epc1} says that for there to be no eventual paths from o to o' these configurations must hold:
\begin{itemize}\setlength\itemsep{0.5em}
\item if there is some object x that o has a path to in \losigma, then x cannot have an eventual path to o'
\item if there is some object x that possesses the capability of o in \losigma, then x cannot have an eventual path to o'
\end{itemize}
Note that object x can refer to the same object as o.
There is one final critical result from \textsc{epc1}.
Notice how, there is a 'recursive' MayAccess$^{Ind,Eve}$(X*,o') in our condition for MayAccess$^{Ind,Eve}$(o,o').
This allows us to recursively expand the condition to incorporate \textit{all} X* intermediate objects in the path leading to o'. This recursive expansion gives us both MayAccess$^{Dir,Now}$ and MayAccess$^{Ind,Now}$ predicates, where the terminating MayAccess$^{Ind,Eve}$(o',o') always returns true if o' exists. This result allows us to define MayAccess$^{Ind,Eve}$(X*,o') in a configuration of paths completely based in \losigma.\\

What about capabilities? Using \textsc{gpec}, we can always expand MayAccess$^{Ind, Now}$(o,X*) into a set of MayAccess$^{Dir, Now}$ predicates that terminate with MayAccess$^{Ind, Now}$(o,o) which is always true if o exists. The final result we will get is a configuration based purely on MayAccess$^{Dir, Now}$ predicates. We expand \textsc{epc1} into \textsc{epc2} to illustrate:

\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Connectivity 2 (EPC2)}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$\{$\linebreak
\ablock \ablock $[$\null
\textcolor{red}{\loexists X$^1$\loin{Obj}. MayAccess$^{Ind,Now}$(o, X$^1$) \loand\linebreak
\ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X$^1$,X*)\linebreak
}
\ablock \ablock \quad \loor \linebreak
\ablock \ablock MayAccess$^{Dir,Now}$(X*,o)$ ]$\linebreak
\ablock \ablock \loand \linebreak
\textcolor{blue}{
\ablock \ablock \loexists Y*\loin Obj.$\{[$MayAccess$^{Ind,Now}$(X*,Y*)\linebreak
\ablock \ablock \ablock \ablock \quad \quad \loor \linebreak
\ablock \ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(Y*,X*)$]$\linebreak
\ablock \ablock \ablock \ablock \quad \loand \linebreak
\ablock \ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(Y*,o')$\}$
}\linebreak
\ablock \ablock $\}$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ \linebreak
\ablock \loforall X*\loin Obj.$\{$\linebreak
\ablock \quad $[$\null
\textcolor{red}{\loforall X$^1$\loin{Obj}. \loneg MayAccess$^{Ind,Now}$(o,X$^1$) \loor\linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X$^1$,X*)\linebreak
}
\ablock \ablock \quad \loand \linebreak
\ablock \quad \loneg MayAccess$^{Dir,Now}$(X*,o)$ ]$\linebreak
\ablock \ablock \loor \linebreak
\textcolor{blue}{
\ablock \loforall Y*\loin Obj. $\{[$\loneg MayAccess$^{Ind,Now}$(X*,Y*)\linebreak
\ablock \ablock \ablock \ablock \quad \loand \linebreak
\ablock \ablock \ablock \quad\: \loneg MayAccess$^{Dir,Now}$(Y*,X*)$]$\linebreak
\ablock \ablock \ablock \quad \loor \linebreak
\ablock \ablock \ablock \quad \loneg MayAccess$^{Ind,Eve}$(Y*,o')$\}$
}\linebreak
\ablock \loimplies\linebreak
	\ablock\loneg MayAccess$^{Ind,Eve}(o,o')$\linebreak
\hrule
\end{logic}

To give a concrete example, let us assume a simple system with only three objects, o, x, o'.
\begin{logic}
\hrule\null
\textsc{\normalsize *Eventual Path Example}\\
M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \ablock $[$\null
\textcolor{red}{MayAccess$^{Ind,Now}$(o,o) \loand\linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(o,x)\linebreak
}
\ablock \ablock \quad \loor \linebreak
\ablock \ablock MayAccess$^{Dir,Now}$(x,o)$ ]$\linebreak
\ablock \ablock \loand \linebreak
\textcolor{blue}{
\ablock \ablock $\{[$
\textcolor{red}{MayAccess$^{Ind,Now}$(x,x) \loand\linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(x,o')\linebreak}
\ablock \ablock \quad \quad \loor \linebreak
\ablock \ablock \quad MayAccess$^{Dir,Now}$(o',x)$]$\linebreak
\ablock \ablock \quad \loand \linebreak
\ablock \ablock \quad MayAccess$^{Ind,Eve}$(o',o')$\}$
}\linebreak
\ablock \ablock $\}$
\linebreak \\
\hrule
\end{logic}
In the example, we make concrete \textsc{epc2} using X* = x and Y* = o', where the expansion terminates at the three predicates, MayAccess$^{Ind,Now}$(o,o), MayAccess$^{Ind,Now}$(x,x) and MayAccess$^{Ind,Eve}$(o',o'). Let us assume o, x, and o' always exists, such that these predicates would return true. This example then illustrates an eventual path from o to o' can only exist if one of these four conditions in \losigma\ hold:
\begin{itemize}\setlength\itemsep{0.5em}
\item o has the capability of x, and x has the capability of o'
\item o has the capability of x, and o' has the capability of x
\item x has the capability of o, and x has the capability of o'
\item o has the capability of x, and x has the capability of o'
\end{itemize}
Note how, these four configurations are present state configurations at \losigma, but allows us to reason about whether a potential path from o to o' can exist in all arising states \losigma' from \losigma.\\

So far we have developed results for paths from o to o', but what about a direct path from o to o'? Luckily, we only need construct \textsc{eec} from \textsc{epc1}, the only difference being we now require the path from o to X* in \textsc{epc1} to be traversable.
